Process and merge data

### 2021-03-20 analysis/01_clean-data.Rmd
Add CIVICUS map data

```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(countrycode)
library(here)
```

@@

# V-Dem
vdem_raw <- read_rds(here("data", "raw_data", "Country_Year_V-Dem_Full+others_R_v10",
                         "V-Dem-CY-Full+Others-v10.rds")) %>% as_tibble()
```

@@

@@

pandem_derog <- derogations_clean %>% 
  left_join(pandem_single, by = c("country_name", "cowcode", "iso3")) %>% 
  left_join(filter(vdem_clean, year == 2019), by = c("country_name", "cowcode")) %>% 
  filter(!is.na(X1)) %>%
  mutate(has_derogation_end = case_when(
    is.na(derogation_end) ~ FALSE,
    str_detect(derogation_end, "N/A") ~ FALSE,
    TRUE ~ TRUE
  )) %>% 
  mutate(has_emergency_end = emlimit == 1) %>% 
  mutate(nonproportional = Type1 > 1 | Type2 == 1 | Type3 > 1) %>% 
  mutate(type = case_when(
    derogate == 1 & has_derogation_end & !nonproportional ~ "A",
    derogate == 1 & (!has_derogation_end | nonproportional) ~ "B",
    derogate == 0 & ratified == 1 & emlaw < 5 ~ "C",
    derogate == 0 & ratified == 1 & emlaw >= 5 ~ "D",
    TRUE ~ "E"
  )) %>% 
  left_join(type_lookup, by = "type")
glimpse(pandem_derog)
```

@@

```{r save-data}
saveRDS(pandem_clean, here("data", "derived_data", "pandem_clean.rds"))
saveRDS(pandem_single, here("data", "derived_data", "pandem_single.rds"))
saveRDS(derogations_clean, here("data", "derived_data", "derogations_clean.rds"))
saveRDS(vdem_clean, here("data", "derived_data", "vdem_clean.rds"))
saveRDS(pandem_derog, here("data", "derived_data", "pandem_derog.rds"))
```

>>>>

```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(countrycode)
library(jsonlite)
library(here)
```

@@

# V-Dem
vdem_raw <- read_rds(here("data", "raw_data", "Country_Year_V-Dem_Full+others_R_v10",
                         "V-Dem-CY-Full+Others-v10.rds")) %>% as_tibble()

# Civicus Monitor
# We downloaded the standalone embeddable widget
# (https://monitor.civicus.org/widgets/world/) as an HTML file with
# `wget https://monitor.civicus.org/widgets/world/` and saved it as index_2021-03-19.html
#
# We then extracted the COUNTRIES_DATA variable embedded in a <script> tag
# (xpath = /html/body/script[5]), which is JSON-ish, but not quite. jsonlite
# can't parse it for whatever reason, but some online JSON formatter and
# validator could, so we ran it through that and saved the resulting clean file
civicus_raw <- read_json(here("data", "raw_data", "Civicus", "civicus_2021-03-19.json"))
```

@@

civicus_lookup <- tribble(
  ~value, ~category,
  1, "Closed",
  2, "Repressed",
  3, "Obstructed",
  4, "Narrowed",
  5, "Open"
) %>% 
  mutate(category = fct_inorder(category, ordered = TRUE))

@@

pandem_derog <- derogations_clean %>% 
  left_join(pandem_single, by = c("country_name", "cowcode", "iso3")) %>% 
  left_join(filter(vdem_clean, year == 2019), by = c("country_name", "cowcode")) %>% 
  filter(!is.na(X1)) %>%
  mutate(has_derogation_end = case_when(
    is.na(derogation_end) ~ FALSE,
    str_detect(derogation_end, "N/A") ~ FALSE,
    TRUE ~ TRUE
  )) %>% 
  mutate(has_emergency_end = emlimit == 1) %>% 
  mutate(nonproportional = Type1 > 1 | Type2 == 1 | Type3 > 1) %>% 
  mutate(type = case_when(
    derogate == 1 & has_derogation_end & !nonproportional ~ "A",
    derogate == 1 & (!has_derogation_end | nonproportional) ~ "B",
    derogate == 0 & ratified == 1 & emlaw < 5 ~ "C",
    derogate == 0 & ratified == 1 & emlaw >= 5 ~ "D",
    TRUE ~ "E"
  )) %>% 
  left_join(type_lookup, by = "type")
glimpse(pandem_derog)

civicus_clean <- civicus_raw %>% 
  as_tibble() %>% 
  slice(1) %>% 
  pivot_longer(everything(), names_to = "name", values_to = "value") %>% 
  mutate(value = map_chr(value, ~.)) %>%
  mutate(value = parse_number(value, na = c("", "NA", "None"))) %>% 
  mutate(country_name = countrycode(name, "iso3c", "country.name",
                                    custom_match = c("KOSOVO" = "XKK",
                                                     "SVT" = "VCT")),
         iso3c = countrycode(country_name, "country.name", "iso3c",
                             custom_match = c("XKK" = "Kosovo",
                                              "VCT" = "Saint Vincent and the Grenadines"))) %>% 
  left_join(civicus_lookup, by = "value") %>% 
  select(-name, -value, -country_name)
glimpse(civicus_clean)
```

@@

```{r save-data}
saveRDS(pandem_clean, here("data", "derived_data", "pandem_clean.rds"))
saveRDS(pandem_single, here("data", "derived_data", "pandem_single.rds"))
saveRDS(derogations_clean, here("data", "derived_data", "derogations_clean.rds"))
saveRDS(vdem_clean, here("data", "derived_data", "vdem_clean.rds"))
saveRDS(pandem_derog, here("data", "derived_data", "pandem_derog.rds"))
saveRDS(civicus_clean, here("data", "derived_data", "civicus_clean.rds"))
```
